---
title: "Coding Assignment 2"
author: "Team 10"
date: "Due: 2022-10-23"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
#Put any packages you need here
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
#Put any packages you need here
library(readr)
library(readxl)
library(car)
```

A Florida health insurance company wants to predict annual claims for
individual clients. The company pulls a random sample of 50 customers.
The owner wishes to charge an actuarially fair premium to ensure a
normal rate of return. The owner collects all of their current
customer's health care expenses from the last year and compares them
with what is known about each customer's plan.

The data on the 50 customers in the sample is as follows:

-   Charges: Total medical expenses for a particular insurance plan (in
    dollars)
-   Age: Age of the primary beneficiary
-   BMI: Primary beneficiary's body mass index (kg/m2)
-   Female: Primary beneficiary's birth sex (0 = Male, 1 = Female)
-   Children: Number of children covered by health insurance plan
    (includes other dependents as well)
-   Smoker: Indicator if primary beneficiary is a smoker (0 =
    non-smoker, 1 = smoker)
-   Cities: Dummy variables for each city with the default being Sanford

Answer the following questions using complete sentences and attach all
output, plots, etc. within this report.

```{r dataset, include=FALSE}
#LM - Load dataset:  Note: removed read.csv and changed to read_csv due to speed of system read
insurance <- read.csv("../Data/insurance_0026_Group10.csv")
```

```{r convert variables}
#LM - Convert all variables into labels
Charges <- insurance[,1]
Age <- insurance[,2]
BMI <- insurance[,3]
Female <- insurance[,4]
Children <- insurance[,5]
Smoker <- insurance[,6]
WinterSprings <- insurance[,7]
WinterPark <- insurance[,8]
Oviedo <- insurance[,9]
```

## Question 1

Generate this regression equation:

$$Charges = \beta_{0}+\beta_{1}*Age+\beta_{2}*BMI+\beta_{3}*Children + \beta_{4}*Female + \beta_{5}*Smoker + \beta_{6}*WinterSprings + \beta_{7}*WinterPark + \beta_{8}*Oviedo$$

```{r Regression Model}
#LM - Regression Equation based on above information
Rmodel <- lm(Charges ~ ., data = insurance )
summary(Rmodel)
```

<font color="#00558b"><b>Question 1 Regression Equation</b></font>

$$Charges = -1806.84 + 165.91*Age + 189.61*BMI - 611.55*Children - 862.22*Female + 19463.54*Smoker - 2696.75*WinterSprings + 1473.58*WinterPark + 740.28*Oviedo$$

Which independent variables test significant in a two-tailed test? Use
$\alpha = 0.2$

<font color="#00558b"><b>Question 1 Analysis</b></font>

<ul>

<li>

Smoker tests significant because      <b>2.3131e-09 \< .2 </b> </br>

<li>

Age tests significant because            <b>0.015478     \< .2</b>

</ul>

## Question 2

Interpret the coefficients of all the independent variables that tested
significant

<font color="#00558b"><b>Question 2 Analysis</b></font>

      <b>Overall</b></br>
      Smoker and Age test statistically significant based on $\alpha = 0.2$ therefore:</br>
      We have sufficient evidence to conclude the slopes of the coefficients are not equal to zero.</br>
      <b>Marginal analysis:  Smoker.  </b></br>
      Dummy variable.            Smoker increases charges by $19463.54.  Non-smoker has no affect on charges.</br>
      <b>Marginal analysis:  Age.</b></br>
      Qualitative variable.      Age charges are $165.91 per year for each year of client's life.

## Question 3

The insurance agent returns with another client:

-   Age: 35
-   BMI: 26
-   Children: 1
-   Gender: Female
-   Smoker: No
-   City: Sanford

Using the regression equation above, predict the amount of medical
expenses associated with this policy. (Provide a 95% confidence interval
as well)

```{r predict}
#LM - From Simple Regression Lecture, added 95% confidence level; copied and pasted from previous assignment
newPrediction <- data.frame  (Age = 35, BMI = 26, Children = 1, Female = 1, Smoker = 0, 
                              WinterSprings = 0, WinterPark = 0, Oviedo = 0)
predict(Rmodel,
        newdata = newPrediction,
        interval = "confidence",
        level = .95)
```

<font color="#00558b"><b>Question 3 Analysis</b></font>

<ul>

<li>

The estimate of charges is $7,456.

<li>

There is a lower limit of $2,409.13 and an upper limit of $12,503.12 at an 95% confidence interval.</b>

</ul>

## Question 4

Compare the model from question 1, to a model that does not include the
group of city dummies. Does including the cities reduce the SSE by a
statistically significant amount? (check the partial F-test for
significance of the dummies as a group). Use $\alpha = 0.05$

```{r delete cities}

#LM - Delete city fields

model_reduced <- insurance[,-(7:9)]

model_reduced <- lm(Charges ~., data = model_reduced)

summary(model_reduced)

```

```{r Regression Model Comparison}
#LM - Regression Equation based on above information
anova(model_reduced, Rmodel)
```

<font color="#00558b"><b>Question 4 Analysis</b></font>

<ul><b>F-Test:</b>
<li>Null hypothesis: Removing the city independent variables from the equation makes no difference:

$$H_{0}: Cities = 0 $$

and the alternative hypothesis indicates removing the cities (independent variables) does make a difference:

$$H_{A}: Cities \neq 0$$

<li>A comparison shows the F statistic to be .8245 and the p-value associated with the F statistic to be .4879.  The P value for F indicates city is not statistically significant.
<li>Since we are given 95% level of significance and since .4879 > .05 we have to accept the null hypothesis as true.
<li>Given the above we therefore conclude that the coefficients are not statistically significant.  Note that including the cities does not reduce the SSE at all, rather it increases the SSE. 

